STJ2012 analysis

### 2016-07-30 analysis/STJ2012_analysis.Rmd
filter samples before OTUS; pairwise spp permanovas

## Filter dataset  
#### Filter OTUs by minimum count
```{r filter_taxa}
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa <- taxa_sums(phy)[which(taxa_sums(phy) >= n)]  
# Remove taxa below threshold count
phy.f <- prune_taxa(names(taxa), phy)  
```

#### Filter samples by minimum count
```{r filter_samples}
# Set threshold number of reads
sn <- 10
# Remove samples with fewer reads than threshold
phy.f <- prune_samples(sample_sums(phy.f)>=sn, phy.f)

@@

@@

@@

>>>>

## Filter dataset
#### Filter samples by minimum count
```{r filter_samples}
# Set threshold number of reads
sn <- 10
# Remove samples with fewer reads than threshold
phy.f <- prune_samples(sample_sums(phy)>=sn, phy)

@@

#### Filter OTUs by minimum count
```{r filter_taxa}
# Set threshold count
n <- 3
# Identify OTUs below threshold count
taxa <- taxa_sums(phy.f)[which(taxa_sums(phy.f) >= n)]  
# Remove taxa below threshold count
phy.f <- prune_taxa(names(taxa), phy.f)  
```

@@

## Differences between species and shores
Whether *Symbiodinium* communities differ among species, or differ between north vs. south shores within species, is evaluated here using permutational analysis of variance (PERMANOVA).

```{r permanovas, cache=T}
# PERMANOVA for differences among species? Post-hoc: which spp are different?

permanova.spp.t <- adonis(phyloseq::distance(phy.f.t, "bray") ~ Species, 
                          data=as(sample_data(phy.f.t), "data.frame"), permutations=9999)

dmat <- as(phyloseq::distance(phy.f.t, "bray"), "matrix")
df <- as(sample_data(phy.f.t), "data.frame")

pairs <- data.frame(t(combn(levels(df$Species), 2)), stringsAsFactors=F)
p.pairs <- setNames(rep(NA, nrow(pairs)), with(pairs, paste(X1, X2, sep='-')))

for (i in 1:nrow(pairs)) {
  dfs <- subset(df, Species %in% c(pairs[i,1], pairs[i,2]))
  dmats <- dmat[rownames(dfs), rownames(dfs)]
  set.seed(152470)
  permanova <- adonis(dmats ~ Species, data=dfs, permutations=999)
  p.pairs[i] <- permanova$aov.tab$`Pr(>F)`[1]
}

multcompLetters(p.pairs)

# Compute within- and between-pool statistics for each Species
shorestats <- data.frame(matrix(dimnames=list(levels(data.frame(sample_data(phy.f.t))$Species), 
                                              c("Species", "n", "overall", "within", 
                                                "between", "R2", "bd", "p")), ncol=8, nrow=10))
for (i in 1:nlevels(data.frame(sample_data(phy.f.t))$Species)) {
  sp <- levels(data.frame(sample_data(phy.f.t))$Species)[i]
  shorestats$Species[i] <- sp
  phy.f.t.species <- subset_samples(phy.f.t, Species==sp)  # Set temp phyloseq object
  shorestats$n[i] <- nsamples(phy.f.t.species)
  md <- meandist(phyloseq::distance(phy.f.t.species, "bray"), 
                 sample_data(phy.f.t.species)$Shore)
  shorestats$within[i] <- summary(md)$W  # Weighted mean dissimilarity within both Pools
  shorestats$between[i] <- summary(md)$B  # Mean dissimilarity between Pools
  shorestats$overall[i] <- summary(md)$D  # Overall dissimilarity
  if (nlevels(as(sample_data(phy.f.t.species), "data.frame")$Shore) > 1) {
    set.seed(70235)
    permanova <- adonis(phyloseq::distance(phy.f.t.species, "bray") ~ Shore, 
                        data=as(sample_data(phy.f.t.species), "data.frame"), permutations=999)
    shorestats$R2[i] <- permanova$aov.tab$"R2"[1]  # PERMANOVA partial R-squared
    shorestats$p[i] <- permanova$aov.tab$"Pr(>F)"[1]  # PERMANOVA p-value
    bd <- betadisper(phyloseq::distance(phy.f.t.species, "bray"),
                     group=as(sample_data(phy.f.t.species), "data.frame")$Shore)
    shorestats$bd[i] <- TukeyHSD(bd)$group[4]
  } else {
    shorestats$R2[i] <- NA
    shorestats$p[i] <- NA
    shorestats$bd[i] <- NA
  }
}
knitr::kable(shorestats)  # dendrogyra, porites furcata, and siderastrea siderea differ by shore
```

@@

## Niche breadth of *Symbiodinium*
```{r nichebreadth}
# Mean dispersion of clades
otubd <- betadisper(d=vegdist(otu_table(phy.f.t), method="bray"), 
                    group=data.frame(tax_table(phy.f.t))$Clade, type="centroid")
otubd

# Number of samples each OTU appears in
nsamp <- data.frame(n=rowSums(otu_table(phy.f.t) != 0))
nsamp$Subtype <- data.frame(tax_table(phy.f.t))[rownames(nsamp), "Subtype"]
nsamp[order(nsamp$n, decreasing=T), ]
nsamp
?order

plot_richness(phy.f, x="Species")
